---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# casewhen
[![Travis build status](https://travis-ci.org/RLesur/casewhen.svg?branch=master)](https://travis-ci.org/RLesur/casewhen) [![Coverage status](https://codecov.io/gh/RLesur/casewhen/branch/master/graph/badge.svg)](https://codecov.io/github/RLesur/casewhen?branch=master) [![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/casewhen)](https://cran.r-project.org/package=casewhen)

The goal of casewhen is to create reusable `dplyr::case_when()` functions.  
`SAS` users may recognise a behavior close to the `SAS FORMATS`.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("RLesur/casewhen")
```

## Motivation  

During data wrangling with `dplyr`, one may use several times identical `case_when()` clauses in different steps. This can lead to a non-DRY code.  
This package provides a convenient mean to define and reuse `dplyr::case_when()` functions.  

## Examples

With `casewhen`, you can easily create reusable `dplyr::case_when()` functions with the function `create_case_when()`:

```{r example1, message=FALSE}
library(dplyr)
library(casewhen)

people <-
  tribble(
    ~name, ~sex, ~seek,
    "Mary", "F", "M",
    "Henry", "M", "F"
  )

cw_sex <- create_case_when(
  x == "F" ~ "Woman",
  x == "M" ~ "Man",
  TRUE ~ as.character(x),
  vars = "x"
)

people %>% 
  mutate(sex_label = cw_sex(sex), 
         seek_label = cw_sex(seek))
```

Reusing a `case_when()` function is mainly convenient when the same transformation is performed across different datasets. 

```{r example2, message=FALSE}
cw_sexyverse <- create_case_when(
  x == "F" | x == "female" & y == "Human" ~ "Woman",
  x == "M" | x == "male" & y == "Human" ~ "Man",
  TRUE ~ as.character(x),
  vars = c("x", "y")
)

people %>% 
  mutate(sex_label = cw_sexyverse(sex, "Human"))

starwars %>%
  mutate(sex_label = cw_sexyverse(gender, species)) %>%
  select(name, gender, species, sex_label) %>%
  head()
```

## `dbplyr` support

You only have to register the `case_when` functions to your connection:

```{r, include=FALSE}
knitr::knit_hooks$set(
  message = function(x, options) {
    x <- sub("<SQL>", "", x)
    x <- gsub("CASE\n", "\n    CASE\n", x, fixed = TRUE)
    x <- gsub("\nWHEN", "\n      WHEN", x, fixed = TRUE)
    x <- gsub("\nEND", "\n    END", x, fixed = TRUE)
    paste0("**`<SQL>`**\n```sql", paste0(x, collapse = "\n"), "```")
  }
)
```


```{r dbplyr, collapse=FALSE, comment=""}
con <- 
  DBI::dbConnect(RSQLite::SQLite(), ":memory:") %>%
  add_case_when(cw_sex, cw_sexyverse)

people_db <- copy_to(con, people)
starwars_db <- copy_to(con, starwars %>% select(name, gender, species))

people_db %>% 
  mutate(sex_label = cw_sex(sex), 
         seek_label = cw_sex(seek)) %>%
  show_query()

people_db %>%
  mutate(sex_label = cw_sexyverse(sex, "Human")) %>%
  show_query()

starwars_db %>%
  mutate(sex_label = cw_sexyverse(gender, species)) %>%
  show_query()

DBI::dbDisconnect(con)
```

